Constraining the Interiors of Asteroids Through Close Encounters

### 2022-04-11 paper/main.tex
Draft 2

% Abstract of the paper
\begin{abstract}
  We investigate the degree to which asteroid interior density distributions can be extracted from rotational velocity data gathered during a close encounter. We derive the equations of motion for an rigid asteroid's orientation and angular velocity to arbitrary precision and use them to generate synthetic rotational velocity data for a representative asteroid on a close Earth encounter. Using Markov Chain Monte Carlo fits, we re-extract the density moments of the asteroid in a wide range of scenarios to measure the degree to which best fit precision is affected. Specifically, we use many injection-retrieval tests to study fit precision's dependence on the asteroid's moment of inertia, observational precision and cadence, orbital parameters, and initial spin pole direction as well as the quantity of near-pericenter data and the central body oblateness. Finally, we discuss the degeneracy between the density moments and the actual density distribution and propose three models to construct a representative density distribution from fit results.
\end{abstract}

@@

Over the past twenty years, the increase in quantity and quality of high-sensitivity, all-sky surveys has prompted the discovery of numerous asteroids. Such advances have been made via ground-based surveys such as the Catalina Sky Survey \cite{larson1998catalina}, Pan-STARRS \cite{kaiser2002pan}, and the Lincoln Near-Earth Asteroid Research project (LINEAR) \cite{stokes2000lincoln} as well as space-based instruments such as he Wide-field Infrared Survey Explorer (WISE) mission \cite{wright2010wide}. Many of these asteroids are relatively small, but some are kilometre-sized and a few are predicted to encounter Earth or other planets closely in the near future. Light curve analysis has been performed on many of these new asteroids to study their surface properties, but analyses that are sensitive to the interior structure of the asteroid are rarer.

Theoretical work has demonstrated that tidal torque alters the angular velocity of an asteroid during a gravitational encounter with another body, and that this tidal torque is sensitive to the interior density distribution of the asteroid \cite{paul88, SCHEERES2000106, ashenberg07, BOUE2009750, HouMar2017}. This two-body generalized system has been studied to different orders and with several different methods. The angular velocity changes induced by tidal torque occur both in asteroid binaries \cite{Naidu_2015, Makarov2022ChaosOO}, and during planetary encounters \cite{RICHARDSON199847, scheeres2004evolution}. Such asteroids need not be tidally disrupted for their rotation state to be perturbed by tidal torque. Some examples of altered rotational velocity have recently been observed, including for the 2013 encounter of (367943) Duende with Earth \cite{MOSKOVITZ2020113519, benson2020spin}, and have been predicted for the 2029 encounter of (99942) Apophis with Earth \cite{SCHEERES2005281, DEMARTINI201993}. The upcoming encounter of Apophis is of particular interest and much attention has been devoted to accurate estimates of its orbital and physical properties \cite{yu2014numerical, hirabayashi2021finite,valvano2022apophis, Lee2022Apophis}, and in particular how its angular velocity will be affected by tidal torque \cite{souchay2014rotational, souchay2018changes}. Altered rotational states have been excluded for 1I/2017 U1 `Oumuamua \cite{KWIECINSKI2018170}.

Such asteroid encounters have further been shown to reveal the physical properties of the encountering asteroid. Refs.~\cite{MOSKOVITZ2020113519, benson2020spin} extract best-fit dimensions for a triaxial ellipsoid model of Duende from light-curve data, and Refs.~\cite{DESCAMPS2020113726} and \cite{BERTHIER2020113990} extract bulk densities and rotational periods for asteroid binaries (3905) Doppler and (617) Patroclus respectively.

More encounter candidates are likely to be discovered in the future --- for Earth or other planets --- potentially by the James Webb Space Telescope (JWST) or Large-aperture Synoptic Survey Telescope (LSST), and general methods for assessing whether an asteroid's encounter properties are conducive to the extraction physical asteroid characteristics from light-curve data is necessary. However, such general statements are not very well studied, since much of previous research has focused on specific encounters.

@@

Throughout this paper, we assume that the asteroid under study is on a hyperbolic encounter with pericenter $r_p$ and excess velocity $v_\infty$. We do not consider any third-body perturbations, and we assume that the body being encountered is much more massive than the asteroid (e.g., Earth or Jupiter).

We make use of two frames of reference to model this system. One is the ``inertial frame,'' with axes denoted by $\unit{X}$, $\unit{Y}$, $\unit{Z}$ and origin placed at the central body's center of mass. $\unit{X}$ points from the central body to the asteroid pericenter and $\unit{Z}$ points parallel to the orbit angular momentum. We will assume that the mass distribution of the central body is known in this inertial frame.

Our second frame is the ``body-fixed'' frame, denoted by $\unit{x}, \unit{y}, \unit{z}$. This frame is fixed with respect to the asteroid's principal axes and rotates with the asteroid, with its origin at the asteroid's center of mass. We will solve for the asteroid's mass distribution with reference to the body-fixed frame. For definiteness, we define $\unit{z}$ to be the principal axis with maximal MOI. In general, we use capital letters to denote vectors in the inertial frame and lowercase vectors to denote vectors in the body-fixed frame.

The relative positions of the body-fixed and inertial frames is given by the position of the asteroid. We represent the relative orientations by $z-y-z$ Euler angles $\alpha$, $\beta$, and $\gamma$, such that a matrix $M$ rotating from the body-fixed to the inertial frame ($M\bm{r} = \bm{R}$) is given by
\begin{equation}
M = R_z(\alpha) R_y(\beta) R_z(\gamma).
\label{eqn:euler-angles}
\end{equation}
Here, $R_i(\theta)$ is a rotation around the unit vector $i$ by $\theta$ (figure \ref{fig:euler-angles}).

\begin{figure}
    \centering
    \begin{tikzpicture}
    \draw[-{Latex[length=3mm]}] (0, 0) -- (-4, 0) node[anchor=east] {$\unit x$};
    \draw[-{Latex[length=3mm]}] (0, 0) -- (2, -3) node[anchor=west] {$\unit y$};
    \draw[-{Latex[length=3mm]}] (0, 0) -- (0, 4) node[anchor=south] {$\unit z$};
    \draw[dashed, -{Latex[length=3mm]}] (0, 0) -- (3.7, -2) node[anchor=north] {$\unit n$};
    \draw[line width=0.5mm,-{Latex[length=3mm]}] (0, 0) -- (-0.5, -3) node[anchor=north] {$\unit X$};
    \draw[line width=0.5mm,-{Latex[length=3mm]}] (0, 0) -- (4, 1) node[anchor=south] {$\unit Y$};
    \draw[line width=0.5mm,-{Latex[length=3mm]}] (0, 0) -- (-1.6, 2.7) node[anchor=south] {$\unit Z$};
    \draw[->] (0.5, -0.75) arc (290:302:2.5);
    \draw (0.9, -0.9) node[anchor=center] {$\alpha$};
    \draw[->] (0, 1.2) arc (130:161:1.3);
    \draw (-0.3, 1.3) node[anchor=center] {$\beta$};
    \draw[->] (0.97, -0.52) arc (330:368:1.3);
    \draw (1.4, -0.25) node[anchor=center] {$\gamma$};
    %\draw[-{Latex[length=3mm]}] (0, 0) -- (0, 4) node[anchor=south] {$\unit Z$};
    %\draw[-{Latex[length=3mm]}] (0, 0) -- (4, 0) node[anchor=west] {$\unit Y$};
    %\draw[-{Latex[length=3mm]}] (0, 0) -- (-2, -2) node[anchor=east] {$\unit X$};
    \end{tikzpicture}
    \caption{$z-y-z$ Euler angles used in this work to express the orientation of the asteroid. Orientation is expressed as a rotation from the body-fixed axes (lowercase) to the inertial axes (bold and uppercase).}
    \label{fig:euler-angles}
\end{figure}

@@

In the next section, it will be shown that only certain aspects of the asteroid density distribution affect tidal torque. We isolate those aspects according to parameters we call ``density moments.'' First, we define the un-normalized spherical harmonics $Y_{\ell m}(\theta, \phi) = P_{\ell m}(\cos \theta)e^{im\phi}$, where $P_{\ell m}$ are the associated Legendre Polynomials without the Condon-Shortley phase. The regular and irregular spherical harmonics then defined as:
\begin{equation}
  \begin{split}
    S_{\ell m}(\bm r) &= (-1)^m (\ell - m)! \frac{Y_{\ell m}(\unit r)}{r^{\ell+1}} \\
    R_{\ell m} (\bm r) &= (-1)^m \frac{r^\ell}{(\ell + m)!} Y_{\ell m}(\unit r).
  \end{split}
\end{equation}
These spherical harmonics obey many useful identities summarized in Ref.~\cite{Gelderen1998TheSO}.

We define the density moments of an asteroid as
\begin{equation}
  K_{\ell m} = \frac{1}{\mu_m a_m^\ell} \int_\mathcal{A} d^3 r \rho_m(\bm r) R_{\ell m}(\bm r).
  \label{eqn:klm}
\end{equation}
Note that these are complex. Here, $\mathcal{A}$ indicates the volume of the asteroid, $\mu_m$ is the mass of the asteroid, $\rho_m(\bm r)$ is the density distribution, and $a_m$ is a real length scale defined (for reasons covered later) as 
\begin{equation}
  a_m^2 = \frac{1}{\mu_m} \int_\mathcal{A} d^3 r \rho(\bm r) r^2.
  \label{eqn:am}
\end{equation}
This length scale can be thought of as akin to the radius of the asteroid (although a spherical, uniform-density asteroid has radius $a_m\sqrt{5/3} $). Both equations \ref{eqn:klm} and \ref{eqn:am} should be computed in the body-fixed axes.

Equations \ref{eqn:klm} and \ref{eqn:am} can be extended to the central body:
\begin{equation}
  \begin{split}
    &J_{\ell m} = \frac{1}{\mu_M a_M^\ell} \int_\mathcal{B} d^3 r \rho_M(\bm r) R_{\ell m}(\bm r)\\
    &a_M^2 = \frac{1}{\mu_M} \int_\mathcal{B} d^3 r \rho_M(\bm r) r^2.
  \label{eqn:jlm}
  \end{split}
\end{equation}
which should be computed in the inertial axes.

Note that both $J_{\ell m}$ and $K_{\ell m}$ are unitless. We call them ``moments'' because the $R_{\ell m}(\bm r)$ contains an $r^\ell$ dependence so that $K_{\ell m}$ is the $\ell$th density moment of the asteroid. The gravitational potential field of the asteroid can be written entirely in terms of $K_{\ell, m}$ and $a_m$, so we expect not to need any information about the density distribution of the asteroid beyond these parameters to compute tidal torque.

@@

By definition, $K_{00}=1$. Furthermore, $K_{1m} = 0$ since the body-fixed frame is centred on the asteroid center of mass. Further calculation reveals that the alignment of the body-fixed frame with the asteroid principal axes also forces $K_{21}=K_{2,-1} = 0$ and $\Im K_{22}=0$. The only physical density moments for $\ell \leq 2$ are therefore $K_{22}$ and $K_{20}$, which are related to the moment of inertia around each principal axis by
\begin{equation}
  \begin{split}
    I_x &= \frac{2}{3}\mu_m a_m^2 \parens{K_{20} - 6 K_{22} + 1}\\
    I_y &= \frac{2}{3}\mu_m a_m^2 \parens{K_{20} + 6 K_{22} + 1}\\
    I_z &= \frac{2}{3}\mu_m a_m^2 \parens{-2K_{20} + 1}.
  \end{split}
  \label{eqn:moi}
\end{equation}
Incidentally, the definition of $a_m$ was chosen to satisfy equation \ref{eqn:moi}.

The physical meaning of $K_{22}$ and $K_{20}$ can also be interpreted via a special case: if the asteroid is a uniform-density triaxial ellipsoid, the moments of inertia are simple to compute in terms of the axis lengths and can be compared to those found in equation \ref{eqn:moi}. This yields semi-axis lengths of 
\begin{equation}
  \begin{split}
  a &= \sqrt{\frac{5}{3}}a_m\sqrt{1-2K_{20}+12K_{22}}\\
  b &= \sqrt{\frac{5}{3}}a_m\sqrt{1-2K_{20}-12K_{22}}\\
  c &= \sqrt{\frac{5}{3}}a_m\sqrt{1+4K_{20}}.
  \label{eqn:ellipsoid-axes}
  \end{split}
\end{equation}

The physical meaning of the higher-order moments $K_{3m}$ can be aided by assessing their symmetry properties. An asteroid that is mirror-symmetric along the $\unit{x}$ axis (meaning $\rho_m(x,y,z)=\rho_m(-x,y,z)$) necessarily sets certain density moments to zero. Which density moments are zeroed by which mirror symmetries is outlined in table \ref{tab:klm-symmetries}. Note that, while no mirror symmetries set $K_{00}$, $K_{20}$, or $K_{22}$ equal to zero, mirror symmetries exist which zero all the other moments, including $K_{3m}$. $\Re K_{32}$, $K_{31}$, and $K_{30}$ are the only $K_{3m}$ components zeroed by only one axis. This will not affect our fit results for $K_{3m}$, but when we compute density moments for sample distributions, we find that most $K_{3m}=0$.

\begin{table}
  \centering
  \begin{tabular}{c|ccccccc}
    \hline
    $\ell$ & $\Re K_{\ell 3}$ & $\Im K_{\ell 3}$ & $\Re K_{\ell 2}$ & $\Im K_{\ell 2}$ & $\Re K_{\ell 1}$ & $\Im K_{\ell 1}$ & $K_{\ell 0}$ \\ \hline
    0 &  &  &  &  &  &  & -\\ 
    1 &  &  &  &  & x & y & z\\ 
    2 &  &  & - & x,y & y,z & x,z & -\\ 
    3 & x,z & y,z & z & x,y,z & x & y & z\\ \hline
  \end{tabular}
  \caption{Axes of mirror symmetry that imply zeroed density moments. For example, for mirror symmetries along $\unit y$ or $\unit z$, $\Im K_{32}=0$. Mirror symmetry along $\unit x$ means $\rho_m(x, y, z) = \rho_m(-x, y, z)$. Dashes indicate that none of the mirror symmetries zero the moment in question. Since $r^2>0$ for $r\neq 0$, no symmetries set $a_m=0$ either.}
  \label{tab:klm-symmetries}
\end{table}

We make one final observation about $K_{\ell m}$: the requirement that $\rho_m(\bm r) \geq 0$ everywhere restricts $K_{\ell m}$. In the case of $K_{2m}$, this fact and the constraint that $I_z$ is larger than $I_x$ or $I_y$ requires $K_{20}$ and $K_{22}$ to fall in the triangle
\begin{equation}
  -\frac{1}{4} \leq K_{20} \leq 0, \qquad |K_{22}| \leq -\frac{K_{20}}{2}.
  \label{eqn:parameter-bounds}
\end{equation}
In practice, we also observe that $|K_{3m}| < 1$; often, $|K_{3m}| < 0.01$ even. For the rest of this paper, we therefore treat an uncertainty on $|K_{3m}|$ of $0.01$ or less as a practical limit on observability, in that uncertainty greater than 0.01 means $K_{3m}$ are effectively unknown. This is only a rough estimate, however.

@@

Derivations for the tidal torque experienced by a rigid body in the gravitational field of a larger mass have been computed by several previous studies \cite{paul88,HouMar2017,BOUE2009750, ashenberg07}, often in terms of the moment of inertia of the rigid body (or higher order moments of inertia), and to varying degrees of precision. A simple, first-order derivation is also easily computable via a power series expansion of the gravitational force of the asteroid, and the result is easily expressed in terms of the asteroid moment of inertia in the inertial frame.

@@

Throughout this paper, we assume that the asteroid remains rigid throughout the encounter. We also assume no third-body perturbations from other Solar System objects. (Actually, third-body perturbing objects are allowed if they are closer to the central body's center of mass than the asteroid perigee distance. Thus, their density moments can be included in the density moments of the central body and this derivation can still be used.) For the sake of simplicity, we also assume that the density moments of the central body are known and do not evolve with time. (I.e., the central body does not rotate).

>>>>

% Abstract of the paper
\begin{abstract}
  We investigate the degree to which asteroid interior density distributions can be extracted from rotational velocity data gathered during a close encounter. We derive the equations of motion for an rigid asteroid's orientation and angular velocity to arbitrary order and use them to generate synthetic rotational velocity data for a representative asteroid on a close Earth encounter. Using Markov Chain Monte Carlo fits, we re-extract the density moments of the asteroid in a wide range of scenarios to measure the degree to which best-fitting parameter precision is affected. Specifically, we use many injection-retrieval tests to study fit precision's dependence on the asteroid's moment of inertia, observational precision and cadence, orbital parameters, and initial spin pole direction as well as the quantity of near-pericentre data and the central body oblateness. Finally, we discuss the degeneracy between the density moments and the actual density distribution and propose three models to construct a representative density distribution from fit results.
\end{abstract}

@@

Over the past twenty years, the increase in quantity and quality of high-sensitivity, all-sky surveys has prompted the discovery of numerous asteroids. Such advances have been made via ground-based surveys such as the Catalina Sky Survey \cite{larson1998catalina}, Pan-STARRS \cite{kaiser2002pan}, and the Lincoln Near-Earth Asteroid Research project (LINEAR) \cite{stokes2000lincoln}, as well as space-based instruments such as the Wide-field Infrared Survey Explorer (WISE) mission \cite{wright2010wide}. Many of these asteroids are relatively small, but some are kilometre-sized and a few are predicted to encounter Earth or other planets closely in the near future. Light curve analysis has been performed on many of these new asteroids to study their surface properties, but analyses that are sensitive to the interior structure of the asteroid are rarer.

Theoretical work has demonstrated that tidal torque alters the angular velocity of an asteroid during a gravitational encounter with another body, and that this tidal torque is sensitive to the interior density distribution of the asteroid \cite{paul88, SCHEERES2000106, ashenberg07, BOUE2009750, HouMar2017}. This two-body generalized system has been studied to different orders and with several different methods. The angular velocity changes induced by tidal torque occur both in asteroid binaries \cite{Naidu_2015, Makarov2022ChaosOO}, and during planetary encounters \cite{RICHARDSON199847, scheeres2004evolution}. Such asteroids need not be tidally disrupted for their rotation state to be perturbed by tidal torque. Some examples of rotational velocity altered during an encounter have recently been observed, including for the 2013 encounter of (367943) Duende with Earth \cite{MOSKOVITZ2020113519, benson2020spin}, and have been predicted for the 2029 encounter of (99942) Apophis with Earth \cite{SCHEERES2005281, DEMARTINI201993}. The upcoming encounter of Apophis is of particular interest and much attention has been devoted to accurate estimates of its orbital and physical properties \cite{yu2014numerical, hirabayashi2021finite,valvano2022apophis, Lee2022Apophis}, and in particular how its angular velocity will be affected by tidal torque \cite{souchay2014rotational, souchay2018changes}. Altered rotational states altered by Solar tidal torque have been excluded for 1I/2017 U1 `Oumuamua \cite{KWIECINSKI2018170}.

Such asteroid encounters have further been shown to reveal the physical properties of the encountering asteroid. Refs.~\cite{MOSKOVITZ2020113519, benson2020spin} extract best-fitting dimensions for a triaxial ellipsoid model of Duende from light-curve data, and Refs.~\cite{DESCAMPS2020113726} and \cite{BERTHIER2020113990} extract bulk densities and rotational periods for asteroid binaries (3905) Doppler and (617) Patroclus respectively.

More encounter candidates are likely to be discovered in the future --- for Earth or other planets --- potentially by the James Webb Space Telescope (JWST) or Large-aperture Synoptic Survey Telescope (LSST), and general methods for assessing whether an asteroid's encounter properties are conducive to the extraction of physical asteroid characteristics from light-curve data is necessary. However, such general encounters are not very well studied, since much of previous research has focused on specific encounters.

@@

Throughout this paper, we assume that the asteroid under study is on a hyperbolic encounter with periapse $r_p$ and excess velocity $v_\infty$. We do not consider any third-body perturbations, and we assume that the body being encountered (the central body, e.g.~a planet) is much more massive than the asteroid.

We make use of two frames of reference to model this system. One is the ``inertial frame,'' with axes denoted by $\unit{X}$, $\unit{Y}$, $\unit{Z}$ and origin placed at the central body's centre of mass and each axis aligned with an asteroid principal axis. $\unit{X}$ points from the central body to the asteroid periapse and $\unit{Z}$ points parallel to the orbit angular momentum. We will assume that the mass distribution of the central body is known in this inertial frame.

Our second frame is the ``body-fixed'' frame, denoted by $\unit{x}, \unit{y}, \unit{z}$. This frame is fixed with respect to the asteroid's principal axes and rotates with the asteroid, with its origin at the asteroid's centre of mass. We will solve for the asteroid's mass distribution with reference to the body-fixed frame. For definiteness, we define $\unit{z}$ to be the principal axis with maximal MOI. In general, we use capital letters to denote vectors in the inertial frame and lowercase vectors to denote vectors in the body-fixed frame.

The difference between the origins of the body-fixed and inertial frames is the position of the asteroid. We represent the relative orientations by $z-y-z$ Euler angles $\alpha$, $\beta$, and $\gamma$, such that a matrix $M$ rotating from the body-fixed to the inertial frame ($M\bm{r} = \bm{R}$) is given by
\begin{equation}
M = R_z(\alpha) R_y(\beta) R_z(\gamma).
\label{eqn:euler-angles}
\end{equation}
Here, $R_i(\theta)$ is a rotation around the unit vector $i$ by $\theta$ (figure \ref{fig:euler-angles}).

\begin{figure}
    \centering
    \begin{tikzpicture}
    \draw[-{Latex[length=3mm]}] (0, 0) -- (-4, 0) node[anchor=east] {$\unit x$};
    \draw[-{Latex[length=3mm]}] (0, 0) -- (2, -3) node[anchor=west] {$\unit y$};
    \draw[-{Latex[length=3mm]}] (0, 0) -- (0, 4) node[anchor=south] {$\unit z$};
    \draw[dashed, -{Latex[length=3mm]}] (0, 0) -- (3.7, -2) node[anchor=north] {$\unit n$};
    \draw[line width=0.5mm,-{Latex[length=3mm]}] (0, 0) -- (-0.5, -3) node[anchor=north] {$\unit X$};
    \draw[line width=0.5mm,-{Latex[length=3mm]}] (0, 0) -- (4, 1) node[anchor=south] {$\unit Y$};
    \draw[line width=0.5mm,-{Latex[length=3mm]}] (0, 0) -- (-1.6, 2.7) node[anchor=south] {$\unit Z$};
    \draw[->] (0.5, -0.75) arc (290:302:2.5);
    \draw (0.9, -0.9) node[anchor=center] {$\alpha$};
    \draw[->] (0, 1.2) arc (130:161:1.3);
    \draw (-0.3, 1.3) node[anchor=center] {$\beta$};
    \draw[->] (0.97, -0.52) arc (330:368:1.3);
    \draw (1.4, -0.25) node[anchor=center] {$\gamma$};
    %\draw[-{Latex[length=3mm]}] (0, 0) -- (0, 4) node[anchor=south] {$\unit Z$};
    %\draw[-{Latex[length=3mm]}] (0, 0) -- (4, 0) node[anchor=west] {$\unit Y$};
    %\draw[-{Latex[length=3mm]}] (0, 0) -- (-2, -2) node[anchor=east] {$\unit X$};
    \end{tikzpicture}
    \caption{$z-y-z$ Euler angles used in this work to express the orientation of the asteroid. Orientation is expressed as a rotation from the body-fixed axes (lowercase) to the inertial axes (bold and uppercase). The origins are co-located for demonstration purposes.}
    \label{fig:euler-angles}
\end{figure}

@@

In the next section, it will be shown that only certain parameters of the asteroid density distribution affect tidal torque called ``density moments.'' First, we define the un-normalized spherical harmonics $Y_{\ell m}(\theta, \phi) = P_{\ell m}(\cos \theta)e^{im\phi}$, where $P_{\ell m}$ are the associated Legendre Polynomials without the Condon-Shortley phase. The regular and irregular spherical harmonics then defined as:
\begin{equation}
  \begin{split}
    S_{\ell m}(\bm r) &= (-1)^m (\ell - m)! \frac{Y_{\ell m}(\unit r)}{r^{\ell+1}} \\
    R_{\ell m} (\bm r) &= (-1)^m \frac{r^\ell}{(\ell + m)!} Y_{\ell m}(\unit r).
  \end{split}
\end{equation}
These spherical harmonics obey many useful identities summarized in Ref.~\cite{Gelderen1998TheSO}, which derives spherical harmonic identities useful for quantum mechanics.

We define the density moments of an asteroid as
\begin{equation}
  K_{\ell m} = \frac{1}{\mu_\mathcal{A} a_\mathcal{A}^\ell} \int_\mathcal{A} d^3 r \rho_\mathcal{A}(\bm r) R_{\ell m}(\bm r).
  \label{eqn:klm}
\end{equation}
Note that these are complex. Here, $\mathcal{A}$ indicates the volume of the asteroid, $\mu_\mathcal{A}$ is the mass of the asteroid, $\rho_\mathcal{A}(\bm r)$ is the density distribution, and $a_\mathcal{A}$ is a real length scale defined (for reasons covered later) as 
\begin{equation}
  a_\mathcal{A}^2 = \frac{1}{\mu_\mathcal{A}} \int_\mathcal{A} d^3 r \rho_\mathcal{A}(\bm r) r^2.
  \label{eqn:am}
\end{equation}
This length scale can be thought of as akin to the radius of the asteroid (although a spherical, uniform-density asteroid has radius $a_\mathcal{A}\sqrt{5/3} $). Both equations \ref{eqn:klm} and \ref{eqn:am} should be computed in the body-fixed axes.

Equations \ref{eqn:klm} and \ref{eqn:am} can be extended to the central body:
\begin{equation}
  \begin{split}
    &J_{\ell m} = \frac{1}{\mu_\mathcal{B} a_\mathcal{B}^\ell} \int_\mathcal{B} d^3 r \rho_\mathcal{B}(\bm r) R_{\ell m}(\bm r)\\
    &a_\mathcal{B}^2 = \frac{1}{\mu_\mathcal{B}} \int_\mathcal{B} d^3 r \rho_\mathcal{B}(\bm r) r^2.
  \end{split}
  \label{eqn:jlm}
\end{equation}
which should be computed in the inertial axes.

Note that both $J_{\ell m}$ and $K_{\ell m}$ are unitless. We call them ``moments'' because the $R_{\ell m}(\bm r)$ contains an $r^\ell$ dependence so that $K_{\ell m}$ is the $\ell$th density moment of the asteroid. The gravitational potential field of the asteroid can be written entirely in terms of $K_{\ell m}$ and $a_\mathcal{A}$, so we expect not to need any information about the density distribution of the asteroid beyond these parameters to compute tidal torque.

@@

By definition, $K_{00}=1$. Furthermore, $K_{1m} = 0$ since the body-fixed frame is centred on the asteroid centre of mass. Further calculation reveals that the alignment of the body-fixed frame with the asteroid principal axes also forces $K_{21}= 0$ and $\Im K_{22}=0$ and the same for $m<0$. The only physical density moments for $\ell \leq 2$ are therefore $K_{22}$ and $K_{20}$, which are related to the moment of inertia around each principal axis by
\begin{equation}
  \begin{split}
    I_x &= \frac{2}{3}\mu_\mathcal{A} a_\mathcal{A}^2 \parens{K_{20} - 6 K_{22} + 1}\\
    I_y &= \frac{2}{3}\mu_\mathcal{A} a_\mathcal{A}^2 \parens{K_{20} + 6 K_{22} + 1}\\
    I_z &= \frac{2}{3}\mu_\mathcal{A} a_\mathcal{A}^2 \parens{-2K_{20} + 1}.
  \end{split}
  \label{eqn:moi}
\end{equation}
Incidentally, the definition of $a_\mathcal{A}$ was chosen to satisfy equation \ref{eqn:moi}.

The physical meaning of $K_{22}$ and $K_{20}$ can also be interpreted via a special case: if the asteroid is a uniform-density triaxial ellipsoid, the moments of inertia are simple to compute in terms of the semi-axis lengths and can be compared to those found in equation \ref{eqn:moi}. This yields semi-axis lengths of 
\begin{equation}
  \begin{split}
  a &= \sqrt{\frac{5}{3}}a_\mathcal{A}\sqrt{1-2K_{20}+12K_{22}}\\
  b &= \sqrt{\frac{5}{3}}a_\mathcal{A}\sqrt{1-2K_{20}-12K_{22}}\\
  c &= \sqrt{\frac{5}{3}}a_\mathcal{A}\sqrt{1+4K_{20}}.
  \label{eqn:ellipsoid-axes}
  \end{split}
\end{equation}

The physical meaning of the higher-order moments $K_{3m}$ can be aided by assessing their symmetry properties. An asteroid that is mirror-symmetric along the $\unit{x}$ axis (meaning $\rho_\mathcal{A}(x,y,z)=\rho_\mathcal{A}(-x,y,z)$) necessarily sets certain density moments to zero. Which density moments are zeroed by which mirror symmetries is outlined in table \ref{tab:klm-symmetries}. Note that, while no mirror symmetries set $K_{00}$, $K_{20}$, or $K_{22}$ equal to zero, mirror symmetries exist which zero all the other moments, including $K_{3m}$. $\Re K_{32}$, $K_{31}$, and $K_{30}$ are the only $K_{3m}$ components zeroed by only one axis. This will not affect our fit results for $K_{3m}$, but when we compute density moments for sample distributions, we find that most $K_{3m}=0$.

\begin{table}
  \centering
  \begin{tabular}{c|ccccccc}
    \hline
    $\ell$ & $\Re K_{\ell 3}$ & $\Im K_{\ell 3}$ & $\Re K_{\ell 2}$ & $\Im K_{\ell 2}$ & $\Re K_{\ell 1}$ & $\Im K_{\ell 1}$ & $K_{\ell 0}$ \\ \hline
    0 &  &  &  &  &  &  & -\\ 
    1 &  &  &  &  & x & y & z\\ 
    2 &  &  & - & x,y & y,z & x,z & -\\ 
    3 & x,z & y,z & z & x,y,z & x & y & z\\ \hline
  \end{tabular}
  \caption{Axes of mirror symmetry that imply zeroed density moments. For example, for mirror symmetries along $\unit y$ or $\unit z$, $\Im K_{32}=0$. Mirror symmetry along $\unit x$ means $\rho_\mathcal{A}(x, y, z) = \rho_\mathcal{A}(-x, y, z)$. Dashes indicate that none of the mirror symmetries zero the moment in question. Since $r^2>0$ for $r\neq 0$, no symmetries set $a_\mathcal{A}=0$ either.}
  \label{tab:klm-symmetries}
\end{table}

Finally, the requirement that $\rho_\mathcal{A}(\bm r) \geq 0$ everywhere restricts $K_{\ell m}$. In the case of $K_{2m}$, this fact and the constraint that $I_z$ is larger than $I_x$ or $I_y$ requires $K_{20}$ and $K_{22}$ to fall in the triangle
\begin{equation}
  -\frac{1}{4} \leq K_{20} \leq 0, \qquad |K_{22}| \leq -\frac{K_{20}}{2}.
  \label{eqn:parameter-bounds}
\end{equation}
In practice, we also observe that $|K_{3m}| < 1$; often, $|K_{3m}| < 0.01$ even. For the rest of this paper, we therefore treat an uncertainty on $|K_{3m}|$ of $0.01$ or less as a practical limit on observability, in that uncertainty greater than 0.01 means $K_{3m}$ are effectively unknown. This is only a rough estimate, however.

@@

Derivations for the tidal torque experienced by a rigid body in the gravitational field of a larger mass have been computed by several previous studies \cite{paul88,HouMar2017,BOUE2009750, ashenberg07}, often in terms of the moment of inertia of the rigid body (or higher order moments of inertia), and to varying degrees of precision. A simple, first-order derivation is also easily computable in terms of the asteroid moment of inertia in the inertial frame.

@@

Throughout this paper, we assume that the asteroid remains rigid throughout the encounter. We also assume no third-body perturbations from other Solar System objects. (Actually, third-body perturbing objects are allowed if they are closer to the central body's centre of mass than the asteroid perigee distance. Then, their density moments can be included in the density moments of the central body and this derivation can still be used.) For the sake of simplicity, we also assume that the density moments of the central body are known and do not evolve with time (i.e., the central body does not rotate).

The gravitational potential energy of the central body is, in its most general form,
\begin{equation}
V(\bm R') = -G\int_\mathcal{B} d^3 R \rho_\mathcal{B}(\bm R) \frac{1}{|\bm{R}-\bm{R'}|}.
\label{eqn:first-pe}
\end{equation}
where $\rho_\mathcal{B}$ is the density distribution of the central body and $\mathcal{B}$ indicates the central body's volume. All vectors here are written in the inertial frame. Given $|\bm{R}| < |\bm{R'}|$, Ref.~\cite{Gelderen1998TheSO} gives the identity
\begin{equation}
  \frac{1}{|\bm R - \bm R'|} = \sum_{\ell, m} R_{\ell m}(\bm R) S_{\ell m}^*(\bm R'),
  \label{eqn:ylm-expansion}
\end{equation}
where the sum is shorthand for $\sum_{\ell, m} = \sum_{\ell = 0}^\infty \sum_{m=-\ell}^\ell$.
We are interested in translating the potential energy of equation \ref{eqn:first-pe} to the body-fixed frame. To do this, we let $\bm{R'} = \bm D + \bm U$, where $\bm D$ is the location of the asteroid in the inertial frame. We further define $\bm U = M\bm u$, where $\bm u$ is in the body-fixed frame and $M$ is the rotation matrix given by the Euler angles $\alpha$, $\beta$, and $\gamma$ (see section \ref{sec:coordinates}). The translation from $\bm {R'}$ to $\bm U$ is then attained by the identity 
\begin{equation}
  S_{\ell m}(\bm R') = \sum_{\ell', m'} (-1)^{\ell'}R^*_{\ell' m'}(\bm U)S_{\ell+\ell', m + m'} (\bm D),
  \label{eqn:ylm-translation}
\end{equation}  
provided by Ref.~\cite{Gelderen1998TheSO}, and from $\bm U$ to $\bm u$ is given by
\begin{equation}
  \begin{split}
    Y_{\ell m}(M\bm u) = \sum_{m'=-\ell}^\ell & (-1)^{m+m'}\sqrt{\frac{(\ell-m')!(\ell+m)!}{(\ell+m')!(\ell-m)!}} \\
    & \times \mathcal{D}^\ell_{mm'}(M)^* Y_{\ell m'}(\bm u).\\
  \end{split}
  \label{eqn:ylm-rotation}
\end{equation}
Here, $\mathcal{D}^\ell_{mm'}(M)$ are the Wigner-$D$ matrices, which are determined by the Euler angles $\alpha$, $\beta$, and $\gamma$ of $M$.

Equations \ref{eqn:first-pe} to \ref{eqn:ylm-rotation} then provide formula for $V(\bm u)$ expressed as a sum of integrals over $\mathcal{B}$ of the central body density $\rho_\mathcal{B}(\bm R)$ times $R_{\ell m}(\bm R)$. These are expressed via equation \ref{eqn:jlm} as $J_{\ell m}$.
